---
title: "The Impact of Urbanization, Culture, and Wealth on Public Art in Toronto"
author: 
  - Alaina Hu
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true
number-sections: true
# bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#install.packages("modelsummary")
library(tidyverse)
library(knitr)
library(dplyr)
library(modelsummary)
library(ggplot2)
```


# Introduction

Public art installations, or art pieces available to the general public in public spaces can be found in many urban cities across the world. Installations can be found in
New York City, Chicago, London, and even Toronto. The city of Toronto has over 400 public art installations provided by two organizations, Public Arts & Monument Collection and the Percent for Public Art Program ("Public Art Map", 2024). Public art not only enhances the urban environment, but it also brings together visitors and residents with local culture and history, creating an educational and bonding experience (Barone, 2023). 

Public art design is heavily integrated with urban planning and development as a whole. It is crucial for overall public art design to consider the overall coordination of the environment, meaning that public art should consider and reflect the environment that it is in (Meng, 2022). Furthermore, public art is also an important symbol of maturity in a city (Liu, 2021). Thus, our research is interested in the factors that influence public art design and selection of location. We analyze the trend between urbanization, cultural diversity, and wealth on the locations of public art installations in Toronto. 

Our research paper begins with the Data section to visualize and further understand the variables we are examining. Next, we go into Discussion of the relationships between the data we are looking for. Finally, we end the paper with the Conclusion section, summarizing the takeaway and future of this research. 



# Data

Data for this research comes from Open Data Toronto, an open source data portal containing various topics of data for the city. 
For the data involved in this paper, we combine `Public Art` and `Ward Profiles (25-Ward Model)`
In order to examine urbanization, cultural diversity, and wealth, we use population, minority population, and average household income as the explanatory variables to represent these factors. To analyze this trend across Toronto, we look at the effect of population, minority population, and average household income on the number of public art installations across the 25 wards of Toronto. 

## Public Art
The Public Art raw dataset includes information such as the source, artist, title, medium, installation year, and ward location on the 409 public art pieces
in Toronto. Since we are only interested in the art pieces and their respective
ward location, we focus on these two aspects of the data. Figure 1, shows the part of the public art data we focus on.

```{r}
#| label: cleaned_art
#| fig-cap: Sample of cleaned public art data
#| echo: false
art_data <- read_csv("/cloud/project/outputs/data/cleaned_art_data.csv", show_col_types = FALSE)

art_data |>
  slice(1:5) |>
  kable(col.names = c("Art ID", "Ward", "Ward Name"))

```

<!-- Talk more about it. -->

<!-- And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.) -->



<!-- Talk way more about it.  -->




<!-- \newpage -->

<!-- \appendix -->

<!-- # Appendix {-} -->


<!-- # Additional data details -->

<!-- # Model details {#sec-model-details} -->

<!-- ## Posterior predictive check -->

<!-- In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows... -->

<!-- In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...  -->

<!-- ```{r} -->
<!-- #| eval: true -->
<!-- #| echo: false -->
<!-- #| message: false -->
<!-- #| warning: false -->
<!-- #| label: fig-ppcheckandposteriorvsprior -->
<!-- #| layout-ncol: 2 -->
<!-- #| fig-cap: "Examining how the model fits, and is affected by, the data" -->
<!-- #| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"] -->

<!-- pp_check(first_model) + -->
<!--   theme_classic() + -->
<!--   theme(legend.position = "bottom") -->

<!-- posterior_vs_prior(first_model) + -->
<!--   theme_minimal() + -->
<!--   scale_color_brewer(palette = "Set1") + -->
<!--   theme(legend.position = "bottom") + -->
<!--   coord_flip() -->
<!-- ``` -->

<!-- ## Diagnostics -->

<!-- @fig-stanareyouokay-1 is a trace plot. It shows... This suggests... -->

<!-- @fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests... -->

<!-- ```{r} -->
<!-- #| echo: false -->
<!-- #| eval: true -->
<!-- #| message: false -->
<!-- #| warning: false -->
<!-- #| label: fig-stanareyouokay -->
<!-- #| fig-cap: "Checking the convergence of the MCMC algorithm" -->
<!-- #| fig-subcap: ["Trace plot", "Rhat plot"] -->
<!-- #| layout-ncol: 2 -->

<!-- plot(first_model, "trace") -->

<!-- plot(first_model, "rhat") -->
<!-- ``` -->



<!-- \newpage -->


# References


